Second Year Paper Summary Statistics

### 2021-11-09 paper/report/summary_stats.Rmd
updating summary stats

```{r packages, results='hide', message=FALSE, echo=F}
library(tidyverse)
library(foreign)
library(stargazer)
```

@@

@@

I am planning to limit my second year project to California and focus on comparing Purple Air PM2.5 data to EPA monitor data. I have not yet finished collecting PM2.5 data from Purple Air monitors but plan to collect years 2018-2021 for all California sensors via the Purple Air API. I am also in the process of collecting EPA PM2.5 data via the EPA API.

@@

>>>>

```{r packages, results='hide', message=FALSE, echo=F}
library(tidyverse)
library(foreign)
library(stargazer)
library(ggplot2)
library(lubridate)
library(kableExtra)
```

@@

```{r summary stats, echo=F, warning=F}
pa = read.csv('../../data/purpleair/date_created_california_manual_request_2021-11-09_metadata.csv') %>%
    mutate(date_start = as.Date(date_start)) %>%
    arrange(date_start) %>% mutate(order=(1:nrow(.))) %>%
    select(-date_created)
pa_counts = pa %>% count(COUNTYFP, year(date_start), sort = TRUE)

year_stats0 = pa %>%
    count(year(date_start))

year_stats1 = pa_counts %>% 
    group_by(`year(date_start)`) %>% 
    summarize(max_n = max(n))
pa_county = read.csv('../../data/purpleair/current_pa_sensors_metadata.csv')
pop = read.csv('../../data/gis/co-est2019-annres-06.xlsx')
aqi = read.csv('../../data/epa/aqi/daily_aqi_by_county_2016-2021.csv')
# stargazer(pa, type='text')

year_stats = pa %>%
    group_by(year(date_start)) %>%
    summarise(pm2.5_ym = mean(pm2.5, na.rm=T)) %>%
    merge(., year_stats0) %>%
    merge(., year_stats1)

year_stats %>%
    rename("Year Sensor Joined the Network"="year(date_start)",
           "Mean PM2.5"="pm2.5_ym",
           "Sensors joined in CA"=n,
           "Max # of Sensors joining in any CA county"=max_n) %>%
    kbl(caption = "Statistics of Purple Air Pollution Sensors Joining the Network", 
        align = 'ccc', digits = 3) %>%
    kable_minimal()

pa2 = merge(pa, pa_county, by.x = "sensor_index", by.y = "id")

```

\FloatBarrier

In the Figure 3 and Table 1, we see that many Purple Air pollution sensors joined
the network in years of high pollution, or just after years of high pollution.
One research question I hope to address is what might explain the spatial and
temporal missingness of pollution data from county-level sensors that the EPA
requires non-attainment counties to keep. Because Purple Air monitors create a 
network spread over the state, I can estimate pollution in areas that don't have
EPA monitors or in places that have EPA monitors but at times when they are shut
off. However, from Figure 2, we can see the distribution of Purple Air monitors
is not distributed evenly or randomly.

@@

<!-- This could become a section of your final paper. Describe the dataset. Provide summary statistics for relevant variables. Discuss issues that may affect your estimates. For example: measurement error, reverse causality, omitted variables bias, weak instruments, failure to converge. -->

<!-- Description: Several page writeup, suitable for inclusion in your final paper. Your advisor must sign off on this output. -->

I am planning to limit my second year project to California and focus on
comparing Purple Air PM2.5 data to EPA monitor data. I have not yet finished
collecting PM2.5 data from Purple Air monitors but plan to collect years
2016-2021 for all California sensors via the Purple Air API. I am also in the
process of collecting EPA PM2.5 data via the EPA API and have a small sample
from Meredith.

@@

# Issues that will affect my estimates of pollution

Measurement error from the Purple Air monitors will need to be addressed. I plan
to explore a combination of averaging over multiple sensors and regression-based
correction using EPA monitors that are near some Purple Air monitors as ground 
truth pollution for that location.

I also plan to estimate a selection model into Purple Air adoption so omitted
variable bias will be a large concern. I will have census data at the census
block group level that I can match to the location of the PA monitors. But that
is not unit-level data so also suffers from measurement error if I am estimating
an adoption model at the unit level. I am also interested in modeling this as
a binary discrete choice model with block- or county-level adoption shares (percentage
of the PA monitors adopted in that month, in that county). With either of these
models of adoption, I will need to instrument PA sensor prices, but since it is
a homogeneous good sold by only one firm and the prices have remained nearly the
same for 4 years, I would predict I will have a weak instrument problem.
